---
title: "Day 4 Loops n Stuff"
author: "Victoria Cutler"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

half_splits <- vector(mode = "numeric", length = (length(quarter_splits)/2))

for (i in seq_along(quarter_splits)) {
half_splits[i] <- quarter_splits[i] + quarter_splits[i+1]
}

```

Make a function to double the user input

```{r}
double_this <- function(my_value) {
  print(2 * my_value)
}

double_this(my_value = 4) # this is more clear to specify the variable and you also don't need to be in order
double_this(400)

double_this(my_value = c(1, 2, 10, 100, 175)) # iterating for us!
```

Write a function that asks for a user's NAME and AGE and returns an exclamation like "I'm name, and i am age years old."

```{r}
name_age_sentence <- function(age, name) {
  print(paste0("hi i'm ", name, ", and i am ", age, " years old."))
}

name_age_sentence(age = 25, name = "victoria")
```

```{r}
humanage_to_animalage <- function(animal, human_age) {
  if (animal == "dog") {
    print(human_age * 7)
  } else if (animal == "goat") {
    print(human_age * 4.7)
  } else {
    print("warning: please enter dog or goat for animal")
  }
}

humanage_to_animalage(animal = "dog", human_age = 12)
humanage_to_animalage(animal = "goat", human_age = 4.2)
humanage_to_animalage(animal = "ladybug", human_age = 2) # this is a silent failure because this ran but nothing showed up
```

```{r}
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)

# Create a function that takes in dog name and outputs "my name is dog name and my favorite food is [food]
dog_menu <- function(name) {
  my_sub <- filter(dog_choice, dog_name == name) ## you always want to ask: is this variable this string? or is this variable this value?
  print(paste0("Hi my name is ", name, " and my favorite food is ", my_sub$food, "!"))
 #  print(my_sub)
}

dog_menu("Khora")
```

```{r}
# essentially we're creating the apply function to calc the mean across columns and print a sentence about it. this can only work over only numeric dfs
means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    col_name <- colnames(df[i])
    print(paste0("The mean value for ", col_name, " is ", col_mean, "."))
  }
}

means_over_columns(df = mtcars)
```

```{r}
clothing <- c("shirt", "pants", "shoes")
color <- c("purple", "red", "blue", "white")

our_clothes <- matrix(nrow = length(color), ncol = length(clothing))

for (i in seq_along(color)) {
  for (j in seq_along(clothing)) {
    clothing_details <- paste(color[i],clothing[j])
    our_clothes[i,j] <- clothing_details
  }
}

our_clothes
```

pseudo code for logistic growth: iterate over multiple times for multiple growth rates:

### Logistic growth

Modeling a population over time and varying growth rate

```{r}
logistic_growth <- function(K, N0, r, time) {
  K / (1 + ((K - N0) / N0) * exp(-r * time))
}

# test:
logistic_growth(K = 5000, N0 = 100, r = 0.07, time = 42)
```

```{r}
time_vec <- seq(from = 0, to = 60, by = .1)
time_vec

pop_35 <- logistic_growth(N0 = 1, K = 8000, r = 0.1, time = time_vec)
pop_35

# we have two separate vectors right now but to plot we want to make these into a data frame thats time by population

pop_time35_df <- data.frame(time_vec, pop_35)

# now we can plot!
pop_time35_plot <- ggplot(pop_time35_df, aes(x = time_vec, y = pop_35)) +
  geom_point()
pop_time35_plot
```

or instead of a vector with things that will iterate our function, we can make a for loop:

```{r}
pop_35_vec <- vector(mode = "numeric", length = length(time_vec))

for (i in seq_along(time_vec)) {
  population <- logistic_growth(N0 = 100, K = 6000, r = 0.27, time = time_vec[i])
  pop_35_vec[i] <- population
}

pop_35_vec
```

Now let's iterate over multiple growth rates with a loop

```{r}
r_seq <- seq(from = 0.2, to = 0.4, by = 0.01)
time_vec <- seq(from = 0, to = 60, by = .1)

out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

for (i in seq_along(r_seq)) {
  for (j in seq_along(time_vec)) {
    population <- logistic_growth(N0 = 100, K = 6000, r = r_seq[i], time = time_vec[j])
    out_matrix[j,i] <- population
  }
}

out_matrix

### now some wrangling to get time as a column on our outmatrix
logistic_df <- data.frame(out_matrix, time = time_vec)

### now we want to name our columns to be meaningful

colnames(logistic_df) <- c(paste0("gr_", r_seq), "time")

### how things are right now, we can't plot growth rate, so let's make that a column as well:

logistic_df_long <- logistic_df |> 
  pivot_longer(cols = -time, values_to = "population", names_to = "growth_rate") ## we want to "pivot longer" aka show all the info for everything except time and the values we're doing that to should be called population and the column names that will now be it's own column is the growth rate

# ^ now we can graph everything all at once!

ggplot(logistic_df_long, aes(x = time, y = population)) +
  geom_line(aes(color = growth_rate))
```

a function with warning and an error message

use `warning` to add a warning

use `stop` to add an error

$$
P = 0.3\rho A v^3
$$

rho: air density (kg / m\^3)

A: area covered by the turbine (m\^2)

v: wind velocity (m/s)

```{r}
calc_windpower <- function(rho, radius, windspeed) {
  
  if (windspeed > 130) {
    warning("Whoa, that's really fast. Are you sure?")
  }
  
  if (rho > 1.225) {
    warning("That air density is suspicious.")
  }
  
  if (radius < 0) {
    stop("Rotor radius must be a positive value in meters")
  }
  
  0.3 * rho * pi * (radius^2) * (windspeed^3)
}

calc_windpower(rho = 1.226, radius = -50, windspeed = 135)
```
